Making Sequential Programs Deterministic

### 2019-04-06 randoop-case-study.tex
Updated annotation counts

In Randoop's 24K LOC of code, we wrote
100 \<@NonDet>,
64 \<@OrderNonDet>,
1391 \<@Det>,
and 109 \<@PolyDet> annotations.

The number of \<@Det> annotations --- one per 17 lines of code --- is much
higher than we would prefer.
Nonetheless, it compares favorably to the
extensive effort by the Randoop developers (\cref{sec:subject-program}),
especially given that that \theDeterminismChecker found issues that the
developers had not.

@@

The second reason is a limitation in \theDeterminismCheckerImplementation,
which is inherited from the Checker Framework~\cite{checker-framework-tool} upon which
\theDeterminismChecker is built.
The Checker Framework does not support qualifier polymorphism over
classes.  This means that fields must be annotated as deterministic.  Our annotations in
Randoop use qualifier polymorphism extensively on methods but cannot use it
on classes, and this leads to many more \<@Det> annotations than would be
needed otherwise.

In future work, we plan to implement qualifier polymorphism for classes,
then re-do the case study.  We estimate this will reduce the number of
\<@Det> annotations needed by about an order of magnitude.

>>>>

In Randoop's 24K LOC of code, we wrote
98 \<@NonDet>,
64 \<@OrderNonDet>,
1083 \<@Det>,
and 103 \<@PolyDet> annotations.

The number of \<@Det> annotations --- one per 22 lines of code --- is much
higher than we would prefer.
Nonetheless, it compares favorably to the
extensive effort by the Randoop developers (\cref{sec:subject-program}),
especially given that that \theDeterminismChecker found issues that the
developers had not.

@@

The second reason is a limitation in \theDeterminismCheckerImplementation,
which is inherited from the Checker Framework~\cite{checker-framework-tool} upon which
\theDeterminismChecker is built.
The Checker Framework does not support qualifier polymorphism over
classes.  This means that fields must be annotated as deterministic.
\<@Det> annotations must be added to methods and constructors that modify these
fields, as well as any methods that call these.
Our annotations in
Randoop use qualifier polymorphism extensively on methods but cannot use it
on classes, and this leads to many more \<@Det> annotations than would be
needed otherwise.

In future work, we plan to implement qualifier polymorphism for classes,
then re-do the case study.  We estimate this will reduce the number of
\<@Det> annotations needed by about an order of magnitude. \todo{Do we plan on
doing other case studies as well?}
